### 2017-04-05 doc/manuscript.Rmd
Added Tony's edits to manuscript

<!--

Outline:

- Primary analysis:
        - Sensitivity analysis with Carb intake?

@@

Word count: 234 / 250

**Objective**: 
Although elevated serum triglyceride (TAG) is a well-described risk factor for
type 2 diabetes, few data are available regarding the role of the specific fatty
acid composition within serum TAG (TAGFA). In this context, our aim was to
examine longitudinal associations of TAGFA with insulin sensitivity (IS) and
beta-cell function.

**Research Design and Methods**:
We used longitudinal data (3 visits over 6 years) from the Prospective
Metabolism and Islet Cell Evaluation (PROMISE) cohort of adults (n=477) at-risk
for diabetes. Glucose and insulin from an OGTT were used to calculate the
Matsuda index (ISI), HOMA2-%S, the Insulinogenic Index over HOMA-IR (IGI/IR),
and the Insulin Secretion-Sensitivity Index-2 (ISSI-2). Gas chromatography
quantified TAGFA composition. Generalized estimating equations (GEE) adjusted for
confounders and partial least squares (PLS) were used for the analysis.

**Results**: 
The outcome variables declined by 14% to 27% over the 6-years. In the adjusted
GEE models with TAGFA as mol%, four TAGFA (14:0, 16:0, 14:1n-7, 16:1n-7) had
strong negative associations while most others (e.g. 18:1n-7, 18:1n-9, 20:2n-6, 
20:5n-3) had strong positive associations. Few associations were seen for 
beta-cell function, except for 16:0 (negative) and 18:1n-7 (positive). PLS
analysis indicated that four TGFA clustered together and strongly predicted
lower IS. These four TAGFA also correlated highly (r>0.4) with clinically
measured TAG.

**Conclusions**:
We found that higher proportions of a cluster of four TAGFA (14:0, 14:1-7,
16:0, 16:1n-7) strongly predicted lower IS as well as hypertriglyceridemia.

>>>>

@@

Word count: 249 / 250

**Objective**: 
Although elevated serum triglyceride (TAG) is a well-described risk factor for
type 2 diabetes (T2DM), few data are available regarding the role of the specific fatty
acid composition within serum TAG (TAGFA) in the pathogenesis of T2DM. Our aim, therefore, was to
examine longitudinal associations of TAGFA with insulin sensitivity (IS) and
beta-cell function.

**Research Design and Methods**:
We used longitudinal data (3 visits over 6 years) from the Prospective
Metabolism and Islet Cell Evaluation (PROMISE) cohort of adults (n=477) who were at-risk
for diabetes at baseline. Glucose and insulin from an OGTT were used to calculate the
Matsuda index (ISI), HOMA2-%S, the Insulinogenic Index over HOMA-IR (IGI/IR),
and the Insulin Secretion-Sensitivity Index-2 (ISSI-2). Gas chromatography
quantified TAGFA composition. Generalized estimating equations (GEE) adjusted for
confounders and partial least squares (PLS) were used for the analysis.

**Results**: 
The outcome variables declined by 14% to 27% over the 6-years. In the adjusted
GEE models, four TAGFA (14:0, 16:0, 14:1n-7, 16:1n-7 as mol%) had
strong negative associations with IS while others (e.g. 18:1n-7, 18:1n-9, 20:2n-6, 
20:5n-3) had strong positive associations. Few associations were seen for 
beta-cell function, except for 16:0 (negative) and 18:1n-7 (positive). PLS
analysis indicated four TAGFA (14:0, 16:0, 14:1n-7, 16:1n-7) that are markers of de novo lipogenesis (DNL)
clustered together and strongly predicted
lower IS. These four TAGFA also correlated highly (r>0.4) with clinically
measured TAG.

**Conclusions**:
We found that higher proportions of a cluster of four DNL TAGFA 
strongly predicted lower IS as well as hypertriglyceridemia.
